---
title: "Footballers' Market Value"
author: "Alexandros Papadopoulos"
date: "2025-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
players_data = read.csv("players_22.csv")
```

## First Chapter
## Data Analysis
Part(i)
```{r}
names(data)

# Keep the necessary columns
players_data = players_data[, c(
  "short_name",
  "player_positions",
  "age",
  "value_eur",
  "wage_eur",
  "height_cm",
  "weight_kg",
  "club_name",
  "league_name",
  "club_joined",
  "club_contract_valid_until",
  "nationality_name",
  "preferred_foot",
  "release_clause_eur",
  "club_position",
  "nation_position",
  "weak_foot",
  "work_rate",
  "international_reputation"
)]


# Rename the columns
names(players_data) = c(
  "Name", "Player_Positions", "Age", "Value_EUR", "Wage_EUR", "Height_cm", "Weight_kg",
  "Club_Name", "League_Name", "Club_Joined", "Contract_Valid_Until",
  "Nationality", "Preferred_Foot", "Release_Clause_EUR",
  "Club_Position", "Nation_Position", "Weak_Foot", "Work_Rate",
  "International_Reputation"
)
```
## Convert positions into numeric codes
Part(ii)
```{r}

# Get the first option as the main position
players_data$Player_Positions = sapply(strsplit(players_data$Player_Positions , ",\\s*"), `[`, 1)

# Function to map position to numeric category
position_to_num <- function(pos) {
  if (pos == "GK") {
    return(1)  # Goalkeeper
  } else if (pos %in% c("CB","RB","LB","RWB","LWB")) {
    return(2)  # Defender
  } else if (pos %in% c("CDM","RDM","LDM","CM","LCM","RCM","CAM","LAM","RAM","LM","RM")) {
    return(3)  # Midfielder
  } else if (pos %in% c("ST","CF","LW","RW","LF","RF")) {
    return(4)  # Forward
  } else {
    return(NA) # Unknown or undefined
  }
}

# Replace the original column directly with numeric codes
players_data$Player_Positions <- sapply(players_data$Player_Positions, position_to_num)

# Now the column itself is numeric
head(players_data[, "Player_Positions"])


```
## Convert Nation names into numeric codes
Part(iii)
```{r}

# Get the FIFA official rankings
fifa_rankings_data = read.csv("fifa_ranking-2024-04-04.csv")

# Convert the rank_date column to Date type if it isn't already
fifa_rankings_data$rank_date = as.Date(fifa_rankings_data$rank_date)

# Keep only rows for 2021-08-12
# Keep only the August update, as FIFA(the game) is released around this time
fifa_rankings_data <- subset(fifa_rankings_data, rank_date == as.Date("2021-08-12"))

# Keep the necessary columns and rename them
fifa_rankings_data = fifa_rankings_data[, c("rank", "country_full", "confederation")]
names(fifa_rankings_data) = c("Rank", "Country", "Confederation")


# # Create a mapping vector
# confederation_map <- c(
#   "CAF" = "Africa",
#   "OFC" = "Oceania",
#   "CONCACAF" = "North America",
#   "UEFA" = "Europe",
#   "AFC" = "Asia",
#   "CONMEBOL" = "South America"
# )
# Create a mapping vector
confederation_map <- c(
  "CAF" = 4,
  "OFC" = 6,
  "CONCACAF" = 3,
  "UEFA" = 1, 
  "AFC" = 5, 
  "CONMEBOL" = 2
)

# Apply the mapping to create a new column
fifa_rankings_data$Confederation <- confederation_map[fifa_rankings_data$Confederation]
names(fifa_rankings_data) = c("Rank", "Country", "Continent")

unique(fifa_rankings_data$Continent)

# Create a named vector: ranking -> continent
continent_map <- setNames(fifa_rankings_data$Continent, fifa_rankings_data$Country)


# Add continent column to your main dataset
players_data$Continent <- continent_map[players_data$Nationality]

# Check
head(players_data[, c("Nationality", "Continent")])


# Create a named vector: country -> ranking
rank_map <- setNames(fifa_rankings_data$Rank, fifa_rankings_data$Country)

# Replace Nationality column with ranking points
players_data$Nationality <- rank_map[players_data$Nationality]

# Check
head(players_data[, "Nationality"])



```

## Convert Clubs into numbers
Part(iv)
```{r}

club_data = read.csv("Soccer_Football Clubs Ranking.csv")
names(club_data)
club_data = club_data[,c("ranking", "club.name")]
names(club_data) = c("Rank", "Club")


# -----------------------------
# Problem:
# The main player dataset (`data`) and the club rankings dataset (`club_data`)
# have slightly different club names. For example, "FC Bayern Munich" vs "Bayern Munich".
# A direct lookup using names would fail and produce NAs.
# -----------------------------

# -----------------------------
# Solution:
# 1. Identify all unique club names in the main dataset.
# 2. For each club, find the most similar name in the rankings dataset using string distance (Jaro-Winkler method).
# 3. Store the matched club name and its ranking in a mapping table.
# 4. Apply this mapping table to the main dataset so that each player receives the correct club ranking.
# -----------------------------

# Benefits:
# - Handles minor differences in club names automatically.
# - Avoids repeated fuzzy matching for every player (only done once per unique club).
# - Ensures consistency: every player from the same club gets the same ranking.
# -----------------------------

# Optional note:
# You could add a threshold on string distance to catch mismatches (e.g., if closest match is too dissimilar, assign NA).

library(stringdist)
library(dplyr)

# Prepare vectors
data_clubs <- unique(players_data$Club_Name)        # all clubs in your main data
club_list <- club_data$Club                 # clubs in ranking data
club_ranks <- setNames(club_data$Rank, club_data$Club)

# Function to find best match
get_best_match <- function(club) {
  distances <- stringdist(tolower(club), tolower(club_list), method = "jw")  # Jaro-Winkler
  best_match <- club_list[which.min(distances)]
  return(best_match)
}

# Create mapping table: Data club -> Closest club in ranking data
mapping <- data.frame(
  Club_Name = data_clubs,
  Closest_Club = sapply(data_clubs, get_best_match),
  stringsAsFactors = FALSE
)

# Add the ranking for each matched club
mapping$Club_Ranking <- club_ranks[mapping$Closest_Club]

# Check mapping table
head(mapping)

# Now apply mapping to the main dataset
players_data <- players_data %>%
  left_join(mapping[, c("Club_Name", "Club_Ranking")], by = "Club_Name")

# Check final result
head(players_data[, c("Club_Name", "Club_Ranking")])



```


## Convert league names into numeric codes
Part(v)(a) # Scrape this website to get the league standing information
```{r}

# Convert league names into numeric codes
unique(players_data$League_Name)

library(rvest)
library(dplyr)
library(readr)

url <- "https://www.kickalgor.com/football-leagues/the-ka-football-leagues-global-rating-for-2021-22-season/the-ka-football-leagues-global-rating-for-2021-22-opening-year-2022-data-table/?utm_source=chatgpt.com"

page <- read_html(url)
tables <- html_table(page, fill = TRUE)

league_data <- tables[[1]]

# Select the correct columns
# Suppose your dataframe is df
league_data <- league_data[, c("Pos", "Country", "LEAGUE")]


# Save as CSV
write_csv(league_data, "league_rankings_2022.csv")

head(league_data)

names(league_data) = c("Rank", "Country", "League")



```

# ---------------------------------------------
# The original file `league_rankings_2022.csv` contained league names
# in mixed formats or languages. To ensure compatibility with the main dataset,
# we translated and standardized all league names into English,
# saving the result as `league_rankings_2022_renamed_english.csv`.

# After translation, we verified that the league names in both datasets 
# (the player dataset and the league rankings dataset) matched correctly.
# Once the names were consistent, we saved the finalized and aligned version 
# as `league_rankings_2022_corrected.csv`.
# This corrected file is now used for mapping league rankings to the player dataset.
# ---------------------------------------------

Part(v)(b)
```{r}

library(dplyr)
library(readr)

# Read your ranking file
league_rankings <- read_csv("league_rankings_2022_corrected.csv")


players_data <- players_data %>%
  left_join(league_rankings %>% dplyr::select(Pos, Unique_Leagues),
            by = c("League_Name" = "Unique_Leagues")) %>%
  mutate(League_Ranking = if_else(is.na(Pos), 67, Pos))

players_data <- dplyr::select(players_data, -Pos)

# Check the result
head(players_data)



```


## Preffered Foot
Part(vi)
```{r}
# Convert Preferred_Foot to numeric: 0 = Left, 1 = Right
players_data$Preferred_Foot <- ifelse(players_data$Preferred_Foot == "Left", 0,
                                  ifelse(players_data$Preferred_Foot == "Right", 1, NA))

# Check
table(players_data$Preferred_Foot)


```

## Contract Length
Part(vii)

```{r}

players_data$Contract_Years_Remained <- players_data$Contract_Valid_Until - 2021
  
```

## Keep only the columns that we need
```{r}

players_data <- subset(players_data, select = -c(Club_Name, League_Name, Club_Joined, Contract_Valid_Until, Club_Position, Nation_Position, Weak_Foot, Work_Rate, International_Reputation, Weight_kg))

# Remove rows with any NA values
players_data <- na.omit(players_data)

write.csv(players_data, "players_data.csv", row.names = FALSE)

names(players_data)


```


## We add another dataset for more data (More parameters)

## Make into a readable form
```{r}
# Read file as raw text
txt <- readLines("2021-2022 Football Player Stats.csv",
                 encoding = "UTF-8")

# Write clean version
writeLines(txt, "2021-2022 Football Player Stats CLEAN.csv",
           useBytes = TRUE)

# Read cleaned file as semicolon-delimited CSV
df <- read.csv("2021-2022 Football Player Stats CLEAN.csv",
               sep = ";",
               encoding = "UTF-8")

write.csv(df, "Stats.csv", row.names = FALSE)


```


### Correct and improve language clarity

```{r}
names(df)


# Keep the necessary columns
df = df[, c(
  "Player", "Pos", "Squad", "Min", "Goals", "PasTotCmp.", "Assists", "Tkl.Int", "Clr", "Err", "Recov"
)]


# Rename the columns
names(df) <- c(
  "Name",
  "Position",
  "Team",
  "Minutes_Played",
  "Goals",
  "Total_Passes_Completed_Perc",
  "Assists",
  "Tackles_Interceptions",
  "Clearances",
  "Errors",
  "Recoveries"
)

df$Name <- iconv(df$Name, from = "latin1", to = "UTF-8")
library(stringi)

df$Name <- stringi::stri_trans_nfc(df$Name)
players_data$Name <- stringi::stri_trans_nfc(players_data$Name)

# Normalize the Name column to NFC (composed) form
df$Name <- stri_trans_nfc(df$Name)
players_data$Name <- stri_trans_nfc(players_data$Name)



```



# Ensure both datasets use the same naming format
#(a)
```{r}
library(dplyr)
library(stringr)

df <- df |>
  mutate(
    Name = Name |>
      str_replace_all("\\?", "") |>
      str_squish() |>
      sapply(function(n) {
        parts <- str_split(n, " +", simplify = TRUE)

        if (length(parts) == 1) {
          # Only one name -> keep it as is
          return(parts[1])
        }

        surname <- parts[1]
        rest    <- paste(parts[-1], collapse = " ")
        paste0(substr(surname, 1, 1), ". ", rest)
      })
  ) |>
  group_by(Name) |>
  slice_max(order_by = Minutes_Played, n = 1) |>   # THIS is the correct column name
  ungroup() |>
  filter(Minutes_Played >= 90)



```

#(b)
```{r}
library(dplyr)
library(stringr)
library(stringi)

# Suffixes, now optional dot
suffixes <- c("Jr\\.?","Sr\\.?")

players_data <- players_data %>%
  mutate(
    Name = Name %>%
      str_replace_all("\\?", "") %>%      # remove question marks
      str_squish() %>%                    # trim extra spaces
      str_remove_all(paste0("\\b(", paste(suffixes, collapse="|"), ")\\b")) %>% # remove Jr, Sr, etc.
      str_squish() %>%
      stri_trans_nfc() %>%                # normalize accents
      sapply(function(n) {
        parts <- str_split(n, " +", simplify = TRUE)
        if(length(parts) == 1) {
          # Single-word name, keep as is
          return(parts)
        } else {
          # First initial + rest
          first <- substr(parts[1], 1, 1)
          rest  <- paste(parts[-1], collapse = " ")
          paste0(first, ". ", rest)
        }
      })
  )


```

# Join the two datasets by the column Name
# This is the final dataset
```{r}
library(dplyr)

final <- df %>%
  inner_join(players_data, by = "Name")

names(final)
#stop("Stopping here because you told me to.")
```

# Modify the columns and save the final dataframe
```{r}
final <- subset(final, select = -c(Name, Position, Team))

# Remove rows with any NA values
final <- na.omit(final)

write.csv(final, "players_data.csv", row.names = FALSE)

names(final)

#stop("Stopping here because you told me to.")

```


# ---------------------------------------------
# ---------------------------------------------
# ---------------------------------------------

## Second Chapter
## Linear Model


```{r}

rm(list = ls()) #clean memory

players_data = read.csv("players_data.csv")

players_data$Value_EUR <- players_data$Value_EUR / 1e6

players_data$Release_Clause_EUR <- players_data$Release_Clause_EUR / 1e6

# Sort by Value_EUR
players_data <- players_data[order(players_data$Value_EUR), ]


players_data <- subset(players_data, select = -c(Release_Clause_EUR, League_Ranking, Preferred_Foot, Errors, Height_cm, Continent))

library(dplyr)

# Keep only players with at least 450 minutes
players_data <- players_data %>%
  filter(Minutes_Played >= 450)

# Keep only players with a value of at least 5 million
players_data <- players_data %>%
  filter(Value_EUR >= 5)


names(players_data)

```
# Correlations
```{r}

# Identify numeric columns except Value_EUR
numeric_cols <- names(players_data)[sapply(players_data, is.numeric)]
numeric_cols <- setdiff(numeric_cols, "Value_EUR")

# Loop over numeric columns and compute correlation
for(col in numeric_cols) {
  corr_value <- cor(players_data$Value_EUR, players_data[[col]], use = "complete.obs")
  cat("Correlation between Value_EUR and", col, "=", round(corr_value, 3), "\n")
}
```




## Part(i)
```{r}

fit <- lm(Value_EUR ~ ., data = players_data)
Yhat <- fit$fitted.values

summary(fit)

```

# for loop
```{r}
# Loop through all predictor columns
predictors <- names(players_data)
predictors <- predictors[predictors != "Value_EUR"]  # exclude dependent variable

for (col in predictors) {
  # Scatter plot
  plot(players_data[[col]], players_data$Value_EUR,
       main = paste("Value_EUR vs", col),
       xlab = col, ylab = "Value_EUR",
       pch = 19, col = "gray")
  
  dataset = players_data[[col]]
  sorted_idx <- order(dataset)
  x_sorted <- dataset[sorted_idx]
  yhat_sorted <- Yhat[sorted_idx]
 
  lines(x_sorted, yhat_sorted, col = 2, lwd = 2)  # lwd=2 makes the line thicker
  
  # Fit simple linear model
  model <- lm(Value_EUR ~ ., data = players_data[, c("Value_EUR", col)])
  
  # Add regression line
  sorted_idx <- order(players_data[[col]])
  x_sorted <- players_data[[col]][sorted_idx]
  yhat_sorted <- model$fitted.values[sorted_idx]
  lines(x_sorted, yhat_sorted, col = "blue", lwd = 2)
}



```

#####
##### Check literature , pws douleuei to fit stin R (An dinei varitita apo mono tou se kapoio variable)


```{r}
plot(players_data$Value_EUR, Yhat)
abline(0,1) 

```




### Errors

```{r}
names(fit)
errors=fit$res
qqnorm(errors)
qqline(errors)

```

Tests for uncorrelated errors with constant variance
```{r}
plot(fit$fitted.values, errors) # den yparhei motivo, fainetai statheri diakymansi
```

Tests for suitability of linear model
```{r}
# Loop through all predictor columns
predictors <- names(players_data)
predictors <- predictors[predictors != "Value_EUR"]  # exclude dependent variable

for (col in predictors) {
  plot(players_data[[col]], errors,
     main = paste("Errors of", col),
     xlab = col, ylab = "Errors",
     pch = 19, col = "gray")
}

```

### Prediction

# Haaland Prediction
```{r}
names(players_data)

# All the stats are per minute
newdata <- data.frame(
  Minutes_Played = 2400,              # total minutes played
  Goals = 1,                         # number of goals
  Total_Passes_Completed_Perc = 75,   # percentage
  Assists = 0.3,                        # number of assists
  Tackles_Interceptions = 0,         # combined tackles + interceptions
  Clearances = 0,
  Recoveries = 1,                     # recoveries
  Player_Positions = 4,      
  Age = 21,                           # numeric encoding of position
  Wage_EUR = 170000,                    # weekly wage
  Nationality = 43,                       # numeric encoding
  Club_Ranking = 19,                     # e.g., league table ranking
  Contract_Years_Remained = 3           # remaining years
)


predict(fit,newdata,interval="conf", level=0.95) #95% diastima empistosynis gia to Y

# 
predict(fit,newdata, interval="pred",level=0.95) #95% diastima provlepsis gia to Y


```


```{r}

newdata <- data.frame(
  Minutes_Played = 2000,
  Goals = 0.3,                       
  Total_Passes_Completed_Perc = 0.91,   
  Assists = 0.5,                       
  Tackles_Interceptions = 4,         
  Clearances = 0,                    
  Recoveries = 2,                    
  Player_Positions = 3,               
  Age = 18,                          
  Wage_EUR = 50000,               
  Nationality = 3,                  
  Club_Ranking = 5,                  
  Contract_Years_Remained = 7        
)

predict(fit,newdata,interval="conf", level=0.95) #95% diastima empistosynis gia to Y

# 
predict(fit,newdata, interval="pred",level=0.95) #95% diastima provlepsis gia to Y
  
```

#Lamine Yamal's Prediction
```{r}

newdata <- data.frame(
  Minutes_Played = 4368,              
  Goals = 0.37,                        
  Total_Passes_Completed_Perc = 82,     
  Assists = 0.52,                      
  Tackles_Interceptions = 1,         
  Clearances = 0,                   
  Recoveries = 1,                     
  Player_Positions = 4,                
  Age = 18,                         
  Wage_EUR = 270540,               
  Nationality = 1,       
  Club_Ranking = 4,                 
  Contract_Years_Remained = 6        
)


predict(fit,newdata,interval="conf", level=0.95) #95% diastima empistosynis gia to Y

predict(fit,newdata, interval="pred",level=0.95) #95% diastima provlepsis gia to Y

```